Accurate binding of calcium to phospholipid bilayers by effective inclusion of electronic polarization

### 2017-10-17 Manuscript/manuscript.tex
Updates to the results and discussion, not ready yet

\begin{table}
  \caption{The relative surface excess and the average number of contacts of Calcium with different
    oxygens in POPC with different simulation models \label{tab:binding} }
  \begin{tabular}{l|c c c}
    model                 & $\Gamma_{Ca}^w$ (nm$^{-2}$)   & $\Sigma ^\mathrm{Ca^{2+}} _\mathrm{PO_4} $ & $\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} $ \\
    \hline
    ECC-lipids            &  $0.07 \pm 0.01 $            &  0.133   &    0.033    \\
    Lipid14/\AA{}qvist    &  $0.13 \pm 0.01 $            &  -       &     -       \\
    Lipid14/Dang          &  $ 0.3 \pm 0.03  $           &  -        &    -       \\
  \end{tabular}
\end{table}
To quantify the binding affinities in different models we calculated 
the relative surface excess from Eq.~\ref{surfexcess}. 
Calcium cations at a molar concentration of $350\,\mathrm{mM}$ provide
$\Gamma_i^w = 0.07 \pm 0.01 \rm{nm}^{-2}$ in the ECC-lipid model, which is significantly
smaller than the values for the Lipid14 model with \AA{}qvist ions, $\Gamma_i^w = 0.13 \pm 0.01 \rm{nm}^{-2}$,
or with Dang ions, $\Gamma_i^w = 0.3 \pm 0.03 \rm{nm}^{-2}$.
\todo{JOE: I'd put the $\Gamma$ values into a table SAMULI: They are now in the table \label{tab:binding} 
with contacts of calcium with different oxygens. However, I am not sure if there is enough information for the
table unless we analyze the contacts for Lipid14 models as well.}
Interestingly, the relative surface excess of 
\ce{NaCl} at $1\,\mathrm{M}$ concentration (ECC-ions~\cite{Pluharova2014}) is qualitatively different from \ce{CaCl2}
having a negative value $\Gamma_i^w = -0.1 \pm 0.01 \rm{nm}^{-2}$, 
meaning that water molecules are preferred to sodium and chloride ions at the membrane-water interface. 
This is in contrast to the most of the available lipid force fields,
which which predict a specific binding of sodium on 
PC lipid bilayers \cite{catte16}.
%\todo{JOE: Add references to NaCl simulation papers that show peaks.
%Add a simple note that there is NBFIX trick in Charmm36 (?).} 
%SAMULI: I think that it is enough to cite NMRlipids II here, which shows all the results
This finding is depicted in Figures~\ref{fig:nacl-dens} and~\ref{fig:cacl-dens},
which show the density profile of ions and water along the membrane normal. 
%The density profiles show
%a larger Ca$^{2+}$ density peak in lipid headgroup region for
%the Lipid14 model with Dang and ECC-ions than for the ECC-lipid model.
\todo{Densities and relative surface excess from NaCl systems to be added.
  The discussion below to be finished after this.}

@@

The contacts between 
%\ce{Ca^{2+}} 
ions and different oxygen moieties of POPC,
%$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O}$,
were analyzed by calculating the probability of 
having the smaller or equal distance than $0.3 \, \mathrm{nm}$, 
as done previously in Ref.~\citenum{javanainen17}.

In contrast to sodium, the dominant contribution
to the binding of \ce{Ca^{2+}} to POPC membranes comes from the phosphate group. 
% Numbers from simulation with 346mM conc. CaCl2
The average number of contacts per lipid 
between \ce{Ca^{2+}} at $290\,\mathrm{mM}$ concentration and any oxygen atom in POPC is 
$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O} = 0.134 $,    % = 17.1/128 
whereas if only phosphate oxygen atoms are considered 
the average number of contacts decreases only by a tiny amount to 
$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{PO_4} = 0.133 $,    % = 17.0/128 . 
and if only carbonyl oxygen atoms are considered 
%(phosphate oxygens contribute, but are not counted as contacts),
this value is merely
$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} = 0.033 $.    % = 4.3/128 . 
% Numbers from simulation with 980/845mM conc. CaCl2
%The average number of contacts per lipid between \ce{Ca^{2+}} and any oxygen atom in POPC is 
%$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O} = 0.264 $,    % = 33.8/128 
%whereas if only phosphate oxygen atoms are considered 
%the average number of contacts decreases only by a tiny amount to 
%$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{PO_4} = 0.262 $,    % = 33.5/128 . 
%and if only carbonyl oxygen atoms are considered 
%(phosphate oxygens contribute, but are not counted as contacts),
%this value is merely
%$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} = 0.066 $.    % = 8.4/128 . 
Hence, calcium is in contact with phosphate oxygens for 99\% of the time it is bound to the membrane, 
whereas only 25\% of the time accounts for contacts with carbonyl oxygens,
and interactions purely with carbonyl oxygens account for less than 1\% of configurations. 
Higher concentration of \ce{CaCl2} naturally increases the amount of contacts per lipid,
for $850\,\mathrm{mM}$ concentration of \ce{CaCl2}
the average number of contacts per lipid is increased to
$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O} = 0.264 $, 
but the distribution of states between phosphate and carbonyl oxygens remains unchanged. 
We corroborate this finding with the probability isodensity contours in Fig.~\ref{fig:volmaps}, 
and density profiles of ions and water along the membrane normal in Fig.~\ref{fig:cacl-dens}.

\begin{figure}[tbp]
  \centering
  \includegraphics[width=6.0cm]{../Fig/volmap_resid10_Ca_Cl_PO4Cent.png} \\
  \includegraphics[width=4.0cm]{../Fig/volmap_resid10_Ca_Cl_sn1Cent.png}
  \includegraphics[width=3.6cm]{../Fig/volmap_resid10_Ca_Cl_sn2Cent.png}
  \caption{\label{fig:volmaps}
      Contours of probability isodensities of \ce{Ca^{2+}} with respect to 
      various moieties fixed in space (highlighted with transparent spheres): 
      phosphate moiety, side chain 1 carbonyl group and side chain 2 carbonyl group.
      Shown contours suggest that the dominant contribution 
      to \ce{Ca^{2+}} binding comes from the phosphate oxygens,
      whereas the interactions with any of the two carbonyl groups are considerably milder.
  }
  \todo{JOE: I'll update this figure with some ensemble of configuration to support binding preference of \ce{Ca^{2+}}}
\end{figure}

and \ce{Na+}
For sodium at $1000\,mM$ concentration 
the average number of contacts per lipid with any oxygen atom in POPC is 
$\Sigma ^\mathrm{Na^{+}} _\mathrm{O} = 0.106 $,    % = 17.1/128 
whereas if only phosphate oxygen atoms are considered 
this number decreases to 
$\Sigma ^\mathrm{Na^{+}} _\mathrm{PO_4} = 0.085 $,    % = 17.0/128 . 
and if only carbonyl oxygen atoms are counted 
it is even lower, 
$\Sigma ^\mathrm{Na^{+}} _\mathrm{O_{carb.}} = 0.048 $.    % = 4.3/128 . 
This means that interaction of sodium and POPC 
has balanced contributions of 55\% from only phosphate oxygens 
(i.e. no contacts with carbonyl oxyens),
and from only carbonyl oxygens with 20\% of configurations 
leaving 25\% for mixed phosphate and carbonyl oxygen configurations.  
This is in line with the calculated negative value of 
relative surface excess of sodium with respect to water, $\Gamma_i^w$, and
their density profiles along the membrane normal (Fig.~\ref{fig:nacl-dens})
suggesting that the populations of bound sodium cations 
naturally decrease with the decreasing density of water.

@@

The residence time of \ce{Ca^{2+}} in a POPC membrane 
is experimentally estimated to be lower than $10\,\mu\mathrm{s}$~\cite{altenbach84}. 
From a recent theoretical work with long simulations \cite{javanainen17}
this time can be roughly estimated 
in the order of $1\,\mu\mathrm{s}$~\cite{javanainen17}. 
Based on Fig. S17 in that work, Ref.~\citenum{javanainen17},
we took the simulation with the apparent highest fluctuations of the number of bound calcium,
which can be obtained at \url{https://doi.org/10.5281/zenodo.259376}
and anaylsed the residence times.
% simulation: (Matti C36 with CaClMK, 800ns total):
We find out that
approximately 20\% of bound residence times of \ce{Ca^{2+}} 
are limited by the length of the simulation, $800\,\mathrm{ns}$, 
and only less than
60\% of bound residence time is smaller than half of the simulation length , $400\,\mathrm{ns}$. 
Other simulations reported in that work show even longer residence times.

This is in contrast to our model, ECC-lipids, 
which estimates the residence time of calcium bound to a POPC membrane to be 
smaller than $60\,\mathrm{ns}$ for 90\% of the time a cation is bound. % exactly $53\,\mathrm{ns}$
i.e. at least two orders of magnitude lower than previous estimates.
In addition, the longest observed residence time in our simulation is $120\,\mathrm{ns}$, 
which is well below the length of our simulations.
This is unfortunately not the case of many theoretical works,
where the residence time cannot be estimated,
or the simulation has not even decorrelated from it's 
initial condition with a population of bound/unbound cations \cite{catte16}, 
because it is limited by the simulation length, 
even though it can be very long (order of $\mu\mathrm{s}$) in some works, e.g. reference~\citenum{javanainen17}. 
Increasing time scales to the order of $\mu\mathrm{s}$ 
doesn't help to alleviate such a problem in current lipid models \cite{javanainen17}. %and also \cite{catte16}, but same simulation was used. 
Such a finding changes the point of view on binding of calcium to PC membranes from
a very strong long-term stable binding with rare exchanges to 
a relatively frequent exchange of clacium cations 
in equilibrium between membrane and solvent.

Interestingly, \ce{Na+} cations exhibit much more rapid exchange between membrane and solvent than calcium.
The residence time of sodium bound to a POPC membrane is
smaller than $2\,\mathrm{ns}$ for 90\% of the time it is bound to the membrane 
with the longest measured bound time being $13\,\mathrm{ns}$. 
Such an exchange between membrane and solvent 
is an order of magnitude faster than for calcium cations. 
A histogram of residence times from our simulations is in Fig.~S\ref{fig:hist_residence_times} in SI. 
\todo{JOE: Plot a histogram of residence times based on our simulations and simulation from \citenum{javanainen17} and put it into SI.
 }

@@

Simulations allow us to directly evaluate the stoichiometry 
by calculating relative propensities of various \ce{Ca^{2+}}:$n\times$POPC complexes
by evaluating contacts between cations and lipids with a cut off radius $0.3\,\mathrm{nm}$. 
The propensities of \ce{Ca^{2+}}:$n\times$POPC complexes at 285~mM concetration 
are in order:
$42\%$ for two lipids,
$30\%$ for one lipid, and
$28\%$ for three lipids.
Apart from a ternary complex, in which one calcium cation binds two lipids, 
we also find complexes with 1~and 3~lipids
occuring with only a slightly lower but similar probability
as shown in Figure~\ref{fig:cacl_complexes}.

\begin{figure}[tbp]
  \centering
  \includegraphics[width=6.0cm]{../Fig/ipython_nb/stoichiometry_NaCl-CaCl2_comparison_Ecc-lipids.pdf} \\
  \caption{\label{fig:cacl_complexes}
      Relative probabilities of existence of \ce{Na^{+}} or \ce{Ca^{2+}} complexes
      with a certain number of POPC lipids. 
      \ce{Na^{+}} complexes were evaluated from the simulation with 1~M concentration;
      and \ce{Ca^{2+}} complexes were evaluated from the simulation with 287~mM concentration.
  }
\end{figure}

Several binding models were proposed and tested~\cite{altenbach84} of which only one,
ternary complex binding model, provided a good fit of the experimental observations. 
In such a model, it is assumed that \ce{Ca^{2+}} cations bind to a POPC membrane 
with a stoichiometry 2~POPC:1~\ce{Ca^{2+}}.
In a later work~\cite{macdonald87}, 
a Langmuir adsorption model (i.e. stoichiometry 1~POPC:1~\ce{Ca^{2+}}) 
was found to provide as good fit as ternary complex model, 
when only low concentrations of \ce{CaCl2} are considered.
Ternary complex model also provides a good fit 
to our simulations with ECC-lipids 
(see Fig.~\ref{fig:cacl-bind} in SI and its caption for details). 
The symmetry of the distribution of complexes from simulation 
as shown in Figure~\ref{fig:cacl_complexes} --   
i.e. almost equal probabilities of complexes with 1 or 3 lipids 
that behave in the total average picture as complexes with 2 lipids -- 
provides clues why ternary complex binding model 
fits both simulation and experimental results relatively well, 
although it is apparently incorrect. 
It is also worth noting that many binding constants in the literature
(e.g., Marsh's handbook \cite{marsh13}) 
are based on the headgroup order parameter changes 
intepreted with the ternary complex binding model.

\todo{SAMULI: I think the we should make clear that the current binding
  constants in, e.g., Marsh's handbook are based on the headgroup order parameter
  changes intepreted with ternary complex binding model. I.e. the experimental raw data
  is excatly the same as we have in this work. Our simulations enable much more versatile interpretation
  of the binding phenomena. I think that this is one the main points of this work, so this comment
  applies to the previous section, introduction, abstract and conclusions as well.
  JOE: agree, this would be a good statement in general. I'll try to put it into the text.
}

Binding of \ce{Na+} cations to a POPC membrane 
mostly follows stoichiomtetry \ce{Na^{+}}:$1\times$POPC 
as shown in Figure~\ref{fig:cacl_complexes}.

>>>>

\begin{table}
  \caption{The relative surface excess and the probabilities of contacts between bound Calcium and different
    oxygen moieties in POPC from different simulation models. The CaCl$_2$ concentration of the ions in water
    is 350mM in the analyzed systems\label{tab:binding} }
  \begin{tabular}{l|c c c}
    model                 & $\Gamma_{Ca}^w$ (nm$^{-2}$)   & $P^\mathrm{Ca^{2+}} _\mathrm{PO_4} $ & $P^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} $ \\
    \hline
    ECC-lipids            &  $0.07 \pm 0.01 $            &  99\%   &    25\%    \\
    Lipid14/\AA{}qvist    &  $0.13 \pm 0.01 $            &  -       &     -       \\
    Lipid14/Dang          &  $ 0.3 \pm 0.03  $           &  -        &    -       \\
  \end{tabular}
\end{table}
To quantify the binding affinities in different models we calculated 
the relative surface excess from Eq.~\ref{surfexcess}. 
Calcium cations at a molar concentration of $350\,\mathrm{mM}$ provide
$\Gamma_i^w = 0.07 \pm 0.01 \rm{nm}^{-2}$ in the ECC-lipid model, which is significantly
smaller than the values for the Lipid14 model with \AA{}qvist ions, $\Gamma_i^w = 0.13 \pm 0.01 \rm{nm}^{-2}$,
or with Dang ions, $\Gamma_i^w = 0.3 \pm 0.03 \rm{nm}^{-2}$.
\todo{JOE: I'd put the $\Gamma$ values into a table SAMULI: They are now in the table \label{tab:binding} 
with contacts of calcium with different oxygens. However, I am not sure if there is enough information for the
table unless we analyze the contacts for Lipid14 models as well.}
Interestingly, the relative surface excess of 
\ce{NaCl} at $1\,\mathrm{M}$ concentration (ECC-ions~\cite{Pluharova2014}) is qualitatively different from \ce{CaCl2}
having a negative value $\Gamma_i^w = -0.1 \pm 0.01 \rm{nm}^{-2}$, 
meaning that water molecules are preferred to sodium and chloride ions at the membrane-water interface. 
This is in contrast to the most of the available lipid force fields,
which which predict a specific binding of sodium on 
PC lipid bilayers \cite{catte16}.
%\todo{JOE: Add references to NaCl simulation papers that show peaks.
%Add a simple note that there is NBFIX trick in Charmm36 (?).} 
%SAMULI: I think that it is enough to cite NMRlipids II here, which shows all the results
This finding is depicted in Figures~\ref{fig:nacl-dens} and~\ref{fig:cacl-dens},
which show the density profile of ions and water along the membrane normal. 
%The density profiles show
%a larger Ca$^{2+}$ density peak in lipid headgroup region for
%the Lipid14 model with Dang and ECC-ions than for the ECC-lipid model.
\todo{Densities and relative surface excess from NaCl systems to be added.
  The discussion below to be finished after this.}

@@

The interactions between 
%\ce{Ca^{2+}} 
ions and different oxygen moieties of POPC,
%$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O}$,
were analyzed by calculating the probability of the
contacts with the smaller or equal distance than $0.3 \, \mathrm{nm}$, 
as done previously in Ref.~\citenum{javanainen17}. 
The results are shown in Table \ref{tab:binding}.
%
%SAMULI: I commented out the average number of contacts. All the essential information
%for this is in the probabilities.
%
%In contrast to sodium, the dominant contribution
%to the binding of \ce{Ca^{2+}} to POPC membranes comes from the phosphate group. 
%% Numbers from simulation with 346mM conc. CaCl2
%The average number of contacts per lipid 
%between \ce{Ca^{2+}} at $290\,\mathrm{mM}$ concentration and any oxygen atom in POPC is 
%$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O} = 0.134 $,    % = 17.1/128 
%whereas if only phosphate oxygen atoms are considered 
%the average number of contacts decreases only by a tiny amount to 
%$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{PO_4} = 0.133 $,    % = 17.0/128 . 
%and if only carbonyl oxygen atoms are considered 
%%(phosphate oxygens contribute, but are not counted as contacts),
%this value is merely
%$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} = 0.033 $.    % = 4.3/128 . 
%
% Numbers from simulation with 980/845mM conc. CaCl2
%The average number of contacts per lipid between \ce{Ca^{2+}} and any oxygen atom in POPC is 
%$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O} = 0.264 $,    % = 33.8/128 
%whereas if only phosphate oxygen atoms are considered 
%the average number of contacts decreases only by a tiny amount to 
%$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{PO_4} = 0.262 $,    % = 33.5/128 . 
%and if only carbonyl oxygen atoms are considered 
%(phosphate oxygens contribute, but are not counted as contacts),
%this value is merely
%$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} = 0.066 $.    % = 8.4/128 . 
The probability for the bound Ca$^{2+}$ ions to be in contact with the phosphate oxygens
is 99\%, indicating the interactions of Calcium ions purely with carbonyl oxygens has the probability of
less than 1\%. However, the probability for the bound Ca$^{2+}$ ion to be in contact with
carbonyl oxygen is 25\%, indicating that roughly quarter of the bound Calcium ions are in close contact
with both, phosphase and carbonyl oxygens. It should be noted that in this analysis carbonyl and
phospate may or may not belong to the same lipid, and the carbonyl chain locate in either of the 
acyl chains. While the higher concentrations of \ce{CaCl2} naturally increased the amount of contacts per lipid,
%for $850\,\mathrm{mM}$ concentration of \ce{CaCl2}
%the average number of contacts per lipid is increased to
%$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O} = 0.264 $, 
the distribution of contacts between phosphate and carbonyl oxygens was not affected.
The most likely contacts between Ca$^{2+}$ ions and phosphate oxygens is visualized with
the probability isodensity contours in Fig.~\ref{fig:volmaps}.
%and density profiles of ions and water along the membrane normal in Fig.~\ref{fig:cacl-dens}.

\begin{figure}[tbp]
  \centering
  \includegraphics[width=6.0cm]{../Fig/volmap_resid10_Ca_Cl_PO4Cent.png} %\\
%  \includegraphics[width=4.0cm]{../Fig/volmap_resid10_Ca_Cl_sn1Cent.png}
%  \includegraphics[width=3.6cm]{../Fig/volmap_resid10_Ca_Cl_sn2Cent.png}
  \caption{\label{fig:volmaps}
      Contours of probability isodensities of \ce{Ca^{2+}} with respect to 
      the phosphate oxygens of POPC from ECC-lipid simulation.
      %moiety, side chain 1 carbonyl group and side chain 2 carbonyl group.
      %Shown contours suggest that the dominant contribution 
      %to \ce{Ca^{2+}} binding comes from the phosphate oxygens,
      %whereas the interactions with any of the two carbonyl groups are considerably milder.
  }
  \todo{JOE: I'll update this figure with some ensemble of configuration to support binding preference of \ce{Ca^{2+}}
  SAMULI: I am not sure if this would be needed anymore.}
\end{figure}

Even thought the \ce{Na+} ions bind to POPC bilayer with the very low probability,
it is possible to analyze the distribution of contacts from simulations with a large
sodium concetration of $1000\,mM$.
%the average number of contacts per lipid with any oxygen atom in POPC is 
%$\Sigma ^\mathrm{Na^{+}} _\mathrm{O} = 0.106 $,    % = 17.1/128 
%whereas if only phosphate oxygen atoms are considered 
%this number decreases to 
%$\Sigma ^\mathrm{Na^{+}} _\mathrm{PO_4} = 0.085 $,    % = 17.0/128 . 
%and if only carbonyl oxygen atoms are counted 
%it is even lower, 
%$\Sigma ^\mathrm{Na^{+}} _\mathrm{O_{carb.}} = 0.048 $.    % = 4.3/128 . 
The results show that the sodium ions bind purely to the phosphate oxygens 
of POPC with the probability of 55\% and purely to the carbonyl oxygens
with the probability of 20\%, leaving the probability of 25\% for the
simultaneous binding to both, phosphate and carbonyl oxygens.  
%This is in line with the calculated negative value of 
%relative surface excess of sodium with respect to water, $\Gamma_i^w$, and
%their density profiles along the membrane normal (Fig.~\ref{fig:nacl-dens})
%suggesting that the populations of bound sodium cations 
%naturally decrease with the decreasing density of water.

In conclusion, the results suggest that the Calcium ions specifically bind to phosphate
oxygens, while occasionally interacting also with carbonyls. This is in good agreement
with the previous conclusions based in large amount of experimental and theoretical
studies \cite{hauser76,hauser78,herbette84,cevc90,binder02}, but slightly updates the
picture given by the previous MD simulation, where the binding to carbonyls was found
to be more pronounced \cite{bockmann03,bockmann04,melcrova16,javanainen17}.
Also the Sodium ions bind most likely to the phosphate oxygens, while the probability
of binding purely to carbonyls is also significant. The physiological relevance
of the details of sodium binding is, however, uncertain due to the very weak binding.

The previous interpretation of the experimental data for the changes of lipid headgroup
order parameters, shown in Fig. \ref{fig:delta_ordPar_CaCl}, concluded that
the experimental data was best explained by the ternary complex binding model, 
where \ce{Ca^{2+}} ion bind to a POPC membrane with a stoichiometry 2~POPC:1~\ce{Ca^{2+}}~\cite{altenbach84}.
However, also a Langmuir adsorption model (i.e. stoichiometry 1~POPC:1~\ce{Ca^{2+}}) 
was successfully fit to the experimental data when only low concentrations of \ce{CaCl2} were considered~\cite{macdonald87}.
These binding models and the measured headgroup order parameter changes
are also inluded in the tabulated experimental binding constants~\cite{marsh13}.

Since the experimentally measured changes of the POPC lipid headgroup order parameters
as a function of CaCl$_2$ concentration are reproduced by the ECC-lipid model,
it can be used to give more versatile interepration of binding stoichiometry than
the simple binding models. The stoichiometry was directly evaluated from simulations
by calculating the relative propensities of various \ce{Ca^{2+}}:$n\times$POPC complexes,
which were defined to occur when cations and lipids were found within a cut off distance of $0.3\,\mathrm{nm}$. 
The observed propensities of
$42\%$ for two lipids,
$30\%$ for one lipid, and
$28\%$ for three lipids 
are shown in Figure~\ref{fig:cacl_complexes} for \ce{Ca^{2+}}:$n\times$POPC complexes at 285~mM concetration.
\todo{Also the larger complexes were analysed, right? So we can write ''Larger complexes were not existent in simulation.'', or something like that?}
While the ternary complex had the largest propensity,
the complexes with 1~and 3~lipids occur with only a slightly lower probability,
suggesting more complex binding stoichiometry than the simple ternary complex model.
%The symmetry of the distribution of complexes,
%from simulation as shown in Figure~\ref{fig:cacl_complexes} --   
%i.e.,
It should be, however, noted that on average one Calcium binds roughly to two lipids,
because the probabilities of forming the complexes with 1 or 3 lipids are almost equal.  
This probably explains the successfull fit of the ternary complex model to the
experimental data. 
\todo{I think that we either have to do the analysis in SI properly,
  i.e., using the real $C_I$ from simulations or remove the sentence below and
the related concent from SI.}
Ternary complex model also provides a good fit 
to our simulations with ECC-lipids 
(see Fig.~\ref{fig:cacl-bind} in SI and its caption for details). 
\begin{figure}[tbp]
  \centering
  \includegraphics[width=6.0cm]{../Fig/ipython_nb/stoichiometry_NaCl-CaCl2_comparison_Ecc-lipids.pdf} \\
  \caption{\label{fig:cacl_complexes}
      Relative probabilities of existence of \ce{Na^{+}} or \ce{Ca^{2+}} complexes
      with a certain number of POPC lipids. 
      \ce{Na^{+}} complexes were evaluated from the simulation with 1~M concentration;
      and \ce{Ca^{2+}} complexes were evaluated from the simulation with 287~mM concentration.
  }
\end{figure}

Also the probabilities of different complexes formed by \ce{Na+} ions and POPC
analyzed from the ECC-lipid model simulation with 1 M concentration of NaCl are
shown in Fig.~\ref{fig:cacl_complexes}. In contrast to Calcium, the Sodium ions
most likely binds to one POPC, while the probability to bind two lipids is less likely
and very small for more than two.
%follows stoichiomtetry \ce{Na^{+}}:$1\times$POPCas shown in

@@

The residence time of \ce{Ca^{2+}} in a POPC membrane 
is experimentally estimated to be lower than $10\,\mu\mathrm{s}$~\cite{altenbach84}.
The recent MD simulations indicate that the equilibration of \ce{Ca^{2+}}
binding to POPC bilayer may take hundreds of nanoseconds~\cite{javanainen17},
which could become an issue for the detailed binding studies with MD simulation
trajectories with the having lenght less than several microseconds \cite{catte16}.

%give a roughly estimate of the order
%of $1\,\mu\mathrm{s}$
%This is unfortunately not the case of many theoretical works,
%where the residence time cannot be estimated,
%or the simulation has not even decorrelated from it's 
%initial condition with a population of bound/unbound cations \cite{catte16}, 
%because it is limited by the simulation length, 
%even though it can be very long (order of $\mu\mathrm{s}$) in some works, e.g. reference~\citenum{javanainen17}. 
%Increasing time scales to the order of $\mu\mathrm{s}$ 
%doesn't help to alleviate such a problem in current lipid models \cite{javanainen17}. %and also \cite{catte16}, but same simulation was used. 
To analyze the residence time in the ECC-lipid model we \todo{describe how the residence time analysis was done}.
Histograms of the residence times analyzed from the ECC-lipid simulation and from the CHARMM36 simulation data
\todo{specify the system, POPC?}
taken from the previous study \cite{javanainen17} (the simulation data was downloaded from Ref.~\citenum{zenodo.259376})
are shown in Fig.~S\ref{fig:hist_residence_times} in SI. 
\todo{JOE: Plot a histogram of residence times based on our simulations and simulation from \citenum{javanainen17} and put it into SI.
 }
%Based on Fig. S17 in that work, Ref.~\citenum{javanainen17},
%we took the simulation with the apparent highest fluctuations of the number of bound calcium,
%which can be obtained at \url{https://doi.org/10.5281/zenodo.259376}
%and anaylsed the residence times.
% simulation: (Matti C36 with CaClMK, 800ns total):

\todo{This paragraph should be clarified, but I cannot do it, because the residence time figure is not there yet.}
In the CHARMM36 simulation
approximately 20\% of bound residence times of \ce{Ca^{2+}} 
are limited by the length of the simulation, $800\,\mathrm{ns}$, 
and only less than
60\% of bound residence time is smaller than half of the simulation length , $400\,\mathrm{ns}$. 
Other simulations with CHARMM36 and Slipids models reported in the previous work show even longer residence times. 
The ECC-lipids model estimates 90\% of the residence times of calcium bound to a POPC membrane to be 
smaller than $60\,\mathrm{ns}$, % exactly $53\,\mathrm{ns}$
which is at least two orders of magnitude lower than for the CHARMM36.
In addition, the longest observed residence time in the ECC-lipid simulation is $120\,\mathrm{ns}$, 
which is well below the length of our simulations.

Interestingly, \ce{Na+} cations exhibit much more rapid exchange between membrane and solvent than calcium.
The residence time of sodium bound to a POPC membrane is
smaller than $2\,\mathrm{ns}$ for 90\% of the time it is bound to the membrane 
with the longest measured bound time being $13\,\mathrm{ns}$. 
Such an exchange between membrane and solvent 
is an order of magnitude faster than for calcium cations.

Such a finding changes the point of view on binding of calcium to PC membranes from
a very strong long-term stable binding with rare exchanges to 
a relatively frequent exchange of clacium cations 
in equilibrium between membrane and solvent. 
The reduced barrier for the Calcium binding is also visible in the density distributions
in Fig.~\ref{fig:cacl-dens}, where the more shallow minimum for the Calcuim
density is observed at the interface.

@@
